# Reproducible Research Assignment 1

## 1. Load and process the data

```{r}
  activity <- read.csv("activity.csv")
```

## 2. The mean total steps taken per day

Subset the data by steps and dates, getting the sum of steps per date.
Create histogram.

```{r}
  steps <- aggregate(steps ~ date, data = activity, na.rm = TRUE, sum)    
  hist(steps$steps, xlab = "Steps per Day", ylab = "Number of Days Steps        Achieved", main = "Number of Steps per Day", col = 5)

```

Mean and median of total steps taken per day.

```{r}
  mean(steps$steps)
  median(steps$steps)
```

## 3.The average daily activity pattern

Subset the data and make line plot.

```{r}
  y <- aggregate(steps ~ interval, data = activity, na.rm = TRUE, mean)
  plot(steps ~ interval, data = y, type = "l", col = 2)
```

Get the interval with the highest average.

```{r}
  y[which.max(y$steps),]
```

## 4. Imputing missing values

Calculate the number of rows with missing values.

```{r}
  sum(rowSums(is.na(activity)))
```

Replace the NA's with average steps from earlier.

```{r}
  for(i in 1:length(activity$steps)){
    if(is.na(activity$steps[i])){
      n <- which(activity$interval[i] == y$interval)
      activity$steps[i] <- y[n,]$steps
    }
  }
```

Histogram of new data, repeating Step 2.

```{r}
  steps <- aggregate(steps ~ date, data = activity, na.rm = TRUE, sum)    
  hist(steps$steps, xlab = "Steps per Day", ylab = "Number of Days Steps        Achieved", main = "Number of Steps per Day", col = 5)
```

Mean and median of new dataset. Only the median difference slightly.

```{r}
  mean(steps$steps)
  median(steps$steps)
```

## 5. Differences in activity patterns on weekends and weekdays

Make sure the date data is in the correct format, then create column with the weekday.

```{r}
  activity$date <- as.Date(activity$date)
  activity$DayofWeek <- weekdays(activity$date)
  
  for(i in 1:length(activity$DayofWeek)){
    if(activity$DayofWeek[i] == "Saturday" | activity$DayofWeek[i] == "Sunday"){
      activity$type[i] <- "Weekend" 
    }
    else
      activity$type[i] <- "Weekday"
  }
  activity$type <- as.factor(activity$type)
  head(activity)
```

Time series plot.

```{r}
  library(ggplot2)
  y <- aggregate(steps ~ interval + type, data = activity, na.rm = TRUE, mean)
  p <- ggplot(y, aes(x = interval, y = steps)) +
    geom_line() +
    facet_grid(type ~ .)
  print(p)
```










